Provisional Title

### 2021-04-28 draft/draft.tex
more methods in paper

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{makecell} % more table customization
\usepackage{mathtools} % nice matrices
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage[table,xcdraw]{xcolor} % text and table background colors
\usepackage[margin=3cm]{geometry} % margins
\usepackage[nottoc,numbib]{tocbibind} % add references to table of contents
\usepackage{lineno} % line numbers

@@

% customization
\setcounter{Maxaffil}{0} % affiliations
\renewcommand\Affilfont{\itshape\small} % style of affiliations text
\makeatletter\renewcommand{\@biblabel}[1]{#1.}\makeatother % change [number] for number. in reference list
\addto{\captionsenglish}{\renewcommand{\bibname}{References}}
\newcommand{\idea}[1]{\textcolor{red}{#1}}
\newcommand{\mr}{\textcolor{red}{\textbf{[missing ref(s)]}}}
\newcommand{\Burl}[1]{\url{#1}}
\zxrsetup{toltxlabel} % cross references: draft & suppl mat
\zexternaldocument*[M-]{draft}
\hypersetup{
  colorlinks = true,
  citecolor=  black,
  linkcolor = {blue},
  filecolor = cyan % controls color of external ref, if used
}
\captionsetup[figure]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Fig.}} % custom figure captions
\captionsetup[table]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Table}} % custom table captions
\hypersetup{citecolor={blue}}
\definecolor{lightgray}{gray}{0.9}

@@

\section{Introduction}\label{intro}

@@

\section{Results}\label{results}

\section{Discussion}\label{discussion}

\section{Methods}\label{methods}

\subsection{Stabilization of environmental communities in simple synthetic environments}
\label{methods:community-assembly}

@@

\subsection{Isolation of dominant species}\label{methods:dominants}

@@

\subsection{Simulations}\label{methods:sim}

@@

\section{Data \& code availability}\label{datacode}

Experimental data and code for the analysis, as well as code for the simulations
and the updated Community Simulator package with instructions to use the
new features can be found in \url{github.com/jdiazc9/coalescence}.

\section{Figures}\label{figs}

>>>>

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{makecell} % more table customization
\usepackage{mathtools} % nice matrices
\usepackage{units} % nice fractions
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage[table,xcdraw]{xcolor} % text and table background colors
\usepackage[margin=3cm]{geometry} % margins
\usepackage[nottoc,numbib]{tocbibind} % add references to table of contents
\usepackage{lineno} % line numbers

@@

% customization
\setcounter{Maxaffil}{0} % affiliations
\renewcommand\Affilfont{\itshape\small} % style of affiliations text
\makeatletter\renewcommand{\@biblabel}[1]{#1.}\makeatother % change [number] for number. in reference list
\addto{\captionsenglish}{\renewcommand{\bibname}{References}}
\newcommand{\idea}[1]{\textcolor{red}{#1}}
\newcommand{\mr}{\textcolor{red}{\textbf{[missing ref(s)]}}}
\newcommand{\Burl}[1]{\url{#1}}
\zxrsetup{toltxlabel} % cross references: draft & suppl mat
\zexternaldocument*[M-]{draft}
\hypersetup{
  colorlinks = true,
  citecolor=  black,
  linkcolor = {blue},
  filecolor = cyan % controls color of external ref, if used
}
\captionsetup[figure]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Figure}} % custom figure captions
\captionsetup[table]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Table}} % custom table captions
\hypersetup{citecolor={blue}}
\definecolor{lightgray}{gray}{0.9}

@@

\section*{Introduction}\label{intro}

@@

\section*{Results \& Discussion}\label{results-discussion}

This is how you refer to the \nameref{intro}.

\section*{Methods}\label{methods}

\subsection*{Stabilization of environmental communities in simple synthetic environments}
\label{methods:community-assembly}

@@

\subsection*{Isolation of dominant species}\label{methods:dominants}

@@

\subsection*{Dominant-dominant and community-community competitions}
\label{methods:competitions}

All possible pairwise dominant-dominant and community-community
competition experiments
were performed by mixing equal volumes (4$\upmu$L) of each of the eight
communities or eight dominants at the onset of the twelfth transfer.
Competitions were set up in their native media,
i.e. in 500$\upmu$L of 1$\times$M9 supplemented with 0.07 C-mol/L of
either glutamine or citrate in 96 deep-well plates.
Plates were incubated at 30$^\circ$C for 48h.
Pairwise competitions were further propagated for seven serial transfers
(roughly 42 generations; Fig. \mr) by transferring 8$\upmu$L of
each culture to fresh media (500$\upmu$L).

\subsection*{Determination of community composition by 16S sequencing}
\label{methods:sequencing}

The sequencing protocol was identical to that described in \cite{Goldford2018}.
Community samples were collected by spinning down at 3500rpm for 25min
in a bench-top centrifuge at room temperature;
cell pellets were stored at -80$^\circ$C before processing.
To maximize Gram-positive bacteria cell wall lysis,
the cell pellets were re-suspended and incubated at 37$^\circ$C for 30min
in enzymatic lysis buffer (20mM Tris-HCl, 2mM sodium EDTA, 1.2\% Triton X-100)
and 20 mg/mL of lysozyme from chicken egg white (Sigma-Aldrich).
After cell lysis, the DNA extraction and purification was performed using the
DNeasy 96 protocol for animal tissues (Qiagen).
The clean DNA in 100$\upmu$L elution buffer of 10mM Tris-HCl, 0.5mM EDTA
at pH 9.0 was quantified using Quan-iT PicoGreen dsDNA Assay Kit
(Molecular Probes, Inc.)
and normalized to 5ng/$\upmu$L in nuclease-free water (Qiagen)
for subsequent 16S rRNA illumina sequencing.
16S rRNA amplicon library preparation was performed following a dual-index
paired-end approach \cite{Kozich2013}.
Briefly, PCR amplicon libraries of V4 regions of the 16S rRNA were prepared 
sing dual-index primers (F515/R805), then pooled and sequenced
using the Illumina MiSeq chemistry and platform.
Each sample went through a 30-cycle PCR in duplicate of 20$\upmu$L
reaction volumes using 5ng of DNA each, dual index primers, and AccuPrime Pfx SuperMix (Invitrogen).
The thermocycling procedure includes a 2min initial denaturation step at
95$^\circ$C, and 30 cycles of the following PCR scheme:
(a) 20 second denaturation at 95$^\circ$C,
(b) 15 second annealing at 55$^\circ$C,
and (c) 5 minute extension at 72$^\circ$C.
The duplicate PCR products of each sample were pooled, purified, and normalized
using SequalPrep PCR cleanup and normalization kit (Invitrogen).
Barcoded amplicon libraries were then pooled and sequenced using
Illumina Miseq v2 reagent kit, which generated 2$\times$250bp paired-end reads
at the Yale Center for Genome Analysis (YCGA).
The sequencing reads were demultiplexed on QIIME 1.9.0 \cite{Caporaso2010}.
The barcodes, indexes, and primers were removed from raw reads,
producing FASTQ files with both the forward and reverse reads for each sample,
ready for DADA2 analysis \cite{Callahan2017}.
DADA2 version 1.1.6 was used to infer unique biological exact sequence variants
(ESVs) for each sample
and na{\"i}ve Bayes was used to assign taxonomy using the SILVA version 123
database \cite{Wang2007,Quast2013}.

\subsection*{Metrics of community distance}
\label{methods:metrics}

Beta-diversity indexes between the invasive and coalesced communities
or the resident and coalesced communities were performed using Bray–Curtis or
Jensen-Shannon similarity metrics. For two abrbitrary communities with
ESV abundances represented by the vectors
$\mathbf{x} = \left( x_1, x_2, \cdots, x_N \right)$
and
$\mathbf{y} = \left( y_1, y_2, \cdots, y_N \right)$
(where $x_i$ and $y_i$ represent the relative abundance of the $i$th ESV in
each community respectively and $N$ is the total number of ESVs), the Bray-Curtis
similarity $\mathrm{BC} \left( \mathbf{x}, \mathbf{y} \right)$ is calculated as
\cite{Bray1957}

\begin{equation}
\mathrm{BC} \left( \mathbf{x}, \mathbf{y} \right) = 
\sum_i \mathrm{min} \left( x_i , y_i \right)
\label{eq:bray-curtis}
\end{equation}
%
and the Jensen-Shannon similarity
$\mathrm{JS} \left( \mathbf{x}, \mathbf{y} \right)$
is defined as one minus the Jensen-Shannon distance (which is, in turn,
the square root of the Jensen-Shannon divergence \cite{Lin1991})

\begin{equation}
\mathrm{JS} \left( \mathbf{x}, \mathbf{y} \right) = 
1 - \sqrt{\frac{1}{2}\mathrm{KL} \left( \mathbf{x}, \mathbf{m} \right) +
\frac{1}{2}\mathrm{KL} \left( \mathbf{y}, \mathbf{m} \right)}
\label{eq:jensen-shannon}
\end{equation}
%
where $\mathbf{m} = \left( \mathbf{x} + \mathbf{y} \right)/2$ and KL denotes the
Kullback-Leibler divergence \cite{Kullback1951}

\begin{equation}
\mathrm{KL} \left( \mathbf{x}, \mathbf{y} \right) = 
\sum_i x_i \; \mathrm{log_2} \left( \frac{x_i}{y_i} \right)
\label{eq:kullback-leibler}
\end{equation}

Additionally, we used alternative metrics that do not account for the specific
fractions of each ESV (that is, only discern whether a ESV is present or absent
in a community). These are the Jaccard similarity
$\mathrm{J} \left( \mathbf{x}, \mathbf{y} \right)$ \cite{Jaccard1912}
and the fraction of the endemic cohort of the invasive community
that persists in the coalesced one,
denoted as
$\mathrm{E} \left( \mathbf{x}_\mathrm{I},
\mathbf{x}_\mathrm{R},
\mathbf{x}_\mathrm{C} \right)$
where the subindices I, R and C correspond to the invasive, resident and coalesced
communities respectively.

\begin{equation}
\mathrm{J} \left( \mathbf{x}, \mathbf{y} \right) = 
\frac{\left|\; \mathbf{x} \cap \mathbf{y} \;\right|}
{\left|\; \mathbf{x} \cup \mathbf{y} \;\right|}
\label{eq:jaccard}
\end{equation}

\begin{equation}
\mathrm{E} \left( \mathbf{x}_\mathrm{I},
\mathbf{x}_\mathrm{R},
\mathbf{x}_\mathrm{C} \right) = 
\left.
\sum_i
f \left( x_{\mathrm{I}i}, x_{\mathrm{R}i}, x_{\mathrm{C}i} \right)
\;
\middle/
\;
\sum_i
g \left( x_{\mathrm{I}i}, x_{\mathrm{R}i} \right)
\right.
\label{eq:endemic}
\end{equation}
%
where we have defined

\begin{equation}
\begin{split}
f \left( x_{\mathrm{I}i}, x_{\mathrm{R}i}, x_{\mathrm{C}i} \right) & =
\begin{cases}
    1 & \text{if } x_{\mathrm{I}i} > 0 \text{ and }
                   x_{\mathrm{R}i} = 0 \text{ and }
                   x_{\mathrm{C}i} > 0 \\
    0 & \text{otherwise}
\end{cases} \\
g \left( x_{\mathrm{I}i}, x_{\mathrm{R}i} \right) & =
\begin{cases}
    1 & \text{if } x_{\mathrm{I}i} > 0 \text{ and }
                   x_{\mathrm{R}i} = 0 \\
    0 & \text{otherwise}
\end{cases}
\end{split}
\label{eq:endemic-f}
\end{equation}

\subsection*{Simulations}\label{methods:sim}

@@

\section*{Data \& code availability}\label{datacode}

Experimental data and code for the analysis, as well as code for the simulations
and the updated Community Simulator package with instructions for the
new features can be found in \url{github.com/jdiazc9/coalescence}.

\section*{Acknowledgements}

The authors wish to thank Joshua Goldford, Pankaj Mehta, Wenping Cui,
Robert Marsland and all members of the Sanchez laboratory
for many helpful discussions.
We also wish to express our gratitude to the Goodman laboratory at Yale
for technical help during the early stages of this project.
The funding for this work partly results from a Scialog Program
sponsored jointly by the Research Corporation for Science Advancement and
the Gordon and Betty Moore Foundation through grants to Yale University by the
Research Corporation and the Simons Foundation.

\clearpage

\section*{Figures}\label{figs}
